Parallel Solving of Multiple Information-Coordinated Global Optimization Problems

### 2021-03-04 2020/JPDC_MCO/2020_JPDC.tex
1.JPDC текст до раздела 4.3

\usepackage{adjustbox}
\usepackage{amsmath}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}

@@

\cite{c1,c2,c3,c4,c5,c6,c7,c8,c9,c10}
\cite{c11,c12,c13,c14,c15,c16,c17,c18,c19,c20}
\cite{c21,c22,c23,c24,c25,c26,c27,c28,c29,c30}
\cite{c31,c32,c33,c34,c35,c36,c37,c38,c39,c40,c41}

@@

The paper proposes a further complication in formulating the problem, when not only one, but several global optimization problems, need to be solved simultaneously. Such problems are encountered, for example, when one needs to find several efficient solutions of multicriteria optimization problems or when part of the variable parameters in the optimization problem can take only a certain number of discrete values. The formulation of such multiple global optimization (MGO) problems is relatively new; for the first time, this type of problem was considered in \cite{c15,c16,c17}. The limited attention given to MGO problems is most likely due to the high computational complexity involved.

In general, MGO problems can be solved by a simple sequential (or parallel) solution of individual problems from a set. However, this approach means the computational complexity of solving problems increases many times. Meanwhile, MGO problems may have a certain informational interrelation -- a similar connectivity is observed, for example, when MGO problems are generated to solve the same multicriteria optimization problem -- see, for example, \cite{c19,c20}. Such interconnectedness of MGO problems can be used to reduce significantly the amount of computation performed when solving the entire set of problems.

The approach proposed in this paper assumes that the computational values of any function of the solvable set of MGO problems can be reduced to the values of any other functions of the set without any time consuming calculations. In such cases, all the search information obtained in solving a particular problem can be used in solving all other problems in the set. Such repeated reuse of search information can significantly reduce the amount of computation performed, up to just a few iterations when solving the next MGO problems. In addition, the presence of informational connectivity significantly expands the possibilities of a parallel solution for MGO problems. The paper proposes two parallel computational schemes: a cooperative scheme, which means the search information between the problems is exchanged when solving several MGO problems in parallel, and a competitive scheme, which occurs when MGO problems solved in parallel compete with each other for the use of available computational resources.

The structure of the paper is as follows. Chapter \ref{sec:2} sets out the problem of multiple global optimization. Chapter \ref{sec:3} presents a general computational scheme for solving global search problems. Chapter \ref{sec:4} proposes parallel methods for solving MGO problems by using computational systems with shared and distributed memory. Chapter \ref{sec:5} contains the results of numerical experiments confirming the effectiveness of the proposed parallel methods. In the conclusion, results are discussed and the main directions for further research are presented.

@@

\begin{equation}\label{eq:2}
D  = \{ y\in R^N: a_i \leq y_i \leq b_i, 1 \leq i \leq N \}
\end{equation}
defined by a set of optimized functions
\begin{equation}\label{eq:3}
F(y) = \{ f_1(y),  f_2(y),\dots, f_s(y) \},
\end{equation}
where $y = (y_1,y_2,\dots,y_N)$ is the vector of varied parameters, $N$ is the dimension of the global optimization problems to be solved, and $D$ is the search domain for the given vectors $a$ and $b$. 
In this paper, it is assumed that the functions $f_i(y)$, $1 \leq i \leq s$, can be multiextremal, and the procedures for calculating the values of these functions at points in the search domain $y \in D$ can be computationally expensive. It is also assumed that the functions $f_i(y)$, $1 \leq i \leq s$, satisfy the Lipschitz condition
\begin{equation}\label{eq:4}
|f_i (y')-f_i (y'')| \leq L_i \|y'-y''\|, y',y''\in D, 1 \leq i \leq s,
\end{equation}
where $L_i$ is the Lipschitz constant for the function $f_i(y)$, $1 \leq i \leq s$, and ${\|*\|}$ denotes the Euclidean norm in $R^N$. Fulfillment of the Lipschitz condition allows us to construct numerical estimates of the possible behavior of minimized functions based on a finite set of calculated function values at points in the search domain.

A key condition of this approach is the assumption that any calculated value of any function $f_i(y)$, $1 \leq i \leq s$, of the set $F$ at an arbitrary point $y \in D$ of the search domain $D$ can be transformed to the value of any other function of the set $F$ at the same point $y \in D$ without any time consuming computations, in a time significantly shorter compared to the computation time values of the functions $f_i(y)$, $1 \leq i \leq s$. In other words, there must be a transformation $Z$ such that

\begin{equation}\label{eq:5}
\begin{matrix}
\forall 1 \leq i, j \leq s, \bar{y} \in D \Rightarrow z_i=Z(z_j ), z_i = f_i (\bar{y}),z_j=f_j (\bar{y}), \\
T(Z) \ll T(f_l (y) ),1 \leq l \leq s,
\end{matrix}
\end{equation}
where $T(*)$ is the computation time of the expression given in parenthesis.

This condition is fulfilled, in part, when solving multicriteria optimization problems (MCO)

\begin{equation}\label{eq:6}
g(y) = (g_1(y), g_2(y), \dots , g_s(y)) \to min,  y\in D.
\end{equation}

To find efficient (unimprovable) solutions, a widely used approach is based on the reduction (scalarization) of the vector criterion $g(y)$ to the general scalar criterion, the minimization of which will lead to the Pareto optimal solution of the MCO problem \cite{c19,c20}. So, when using mini-max convolution, the vector criterion $g(y)$ from (\ref{eq:6}) reduces to the scalar criterion
\begin{equation}\label{eq:7}
\begin{matrix}
G(\lambda,y)=\max{(\lambda_i g_i(y),1 \leq i \leq s)},	\\
\lambda=(\lambda_1,\lambda_2,\dots,\lambda_s)\in \Lambda \subset R^s:\sum_{i=1}^s{\lambda_i =1},\lambda_i\geq 0,1 \leq i \leq s.
\end{matrix}
\end{equation}

If it is necessary to find several efficient decisions to the MCO problem (\ref{eq:6}), the optimization of the scalar criterion $G(\lambda,y)$ from (\ref{eq:7}) is performed for different values of the coefficients $\lambda \in \Lambda$ of the convolution of the criteria $g_i(y)$, $1 \leq i \leq s$. Thus, the problem of finding a set of efficient decisions for the MCO problem can be represented as a statement of the MGO problem (\ref{eq:1})-(\ref{eq:3}), where the optimized functions $f_i(y)$, $1 \leq i \leq s$, and the sets $F$ are defined as
\begin{equation}\label{eq:8}
f_i (y)= G(\lambda_i,y),1 \leq i \leq s.
\end{equation}

Thus, if the values of efficiency criteria $g_i(y)$, $1 \leq i \leq s$, computed at some point $y=\bar{y}$ of the search domain $D$, the value of the scalar criterion $G(\lambda,y)$ (and, accordingly, the values of the functions $f_i(y)$, $1 \leq i \leq s$, from a set $F$) can be computed at the point $\bar{y} \in D$ for any values of the coefficients $\lambda \in \Lambda$ of the convolution $g_i(y)$, $1 \leq i \leq s$ (the transformation of $Z$ from (\ref{eq:5}) is determined by the expression (\ref{eq:7}) at fixed values of the criteria $g_i(y)$, $1 \leq i \leq s$) \cite{c15,c16,c17}.

The ability to quickly convert the calculated values of optimized functions determines the need to accumulate and use of search information obtained in the calculation process

\begin{equation}\label{eq:9}
\omega_k (j)=\{(y^i,f^i=f_j(y^i)):1 \leq i \leq k \},1 \leq j \leq s,
\end{equation}
where $j$, $1 \leq j \leq s$, is the number of the current minimized function $f_j(y)$ of the set $F$ from (\ref{eq:3}), $k$ is the number of iterations of the global search, $y^i \in D$, $1 \leq i \leq k$,  are points in which the values of the optimized problems of the set $F$ were computed, and $z^i$, $1 \leq i \leq k$, are the values of the function $f_j(y)$ at these points. Global optimization methods can use this search information, thereby ensuring the ability to solve each successive problem of the $F$ set, taking into account the results of all previous calculations. Numerical experiments that were performed (see Section \ref{sec:5}), confirm that such repeated use of search information reduces the number of iterations of the global search required to solve the next problem of the set $F$ (even to the point of performing just a few search iterations).

In conclusion, it should be noted that, in general, the set of minimized functions $F$ from (\ref{eq:3}) can be formed dynamically in the course of calculations by adding new or removing existing global optimization functions

\begin{equation}\label{eq:10}
F'=F+(-) f(y).
\end{equation}

@@

The approach proposed in this paper is based on two key ideas used in the framework of the information-statistical theory of global optimization \cite{c5}.

In accordance with this theory, first of all, dimension reduction is achieved using the Peano space-filling curves (evolvents) $y(x)$, uniquely and continuously mapping the interval $[0,1]$ to the $N$-dimensional domain $D$ -- see, for example, \cite{c5,c21}. As a result of this reduction, multidimensional global optimization problems (\ref{eq:7}) can be reduced to one-dimensional ones:

@@

Dimensionality reduction significantly decreases the computational complexity of analyzing multidimensional data to highlight the subdomains of the search domain $D$ for the efficient continuation of the global search, and reduces the severity of the ``curse of dimensionality'' problem in solving multidimensional optimization problems. After dimensionality reduction, the search information $\omega_k$ from (\ref{eq:9}) takes the form

\begin{equation}\label{eq:13}
A_k (j)=\{ (x_i, y(x_i), z_i=f_j (y(x_i) ): 1 \leq i \leq k,1 \leq j \leq s \},
\end{equation}
in which, in addition to the reduced representation, the data is arranged in the ascending order\footnote{Data ordering is reflected by using a subscript.} of points $x_i$,  $1 \leq i \leq k$, i.e.

\begin{equation}\label{eq:14}
x_1< x_2< \cdots < x_k
\end{equation}
for a more efficient execution of global search algorithms. It should be noted that in optimizing the reduced one-dimensional functions $f_i (y(x))$, $1 \leq i \leq s$, many one-dimensional global search algorithms can be used (possibly after some generalization).

Numerical methods for constructing approximations of Peano space-filling curves (evolvents) with a given accuracy for a given dimensionality are considered in \cite{c5}; examples of such approximations are shown in Fig. \ref{fig:1}.

@@

Further, in the framework of information-statistical theory, a general computational scheme of global optimization algorithms is proposed. This scheme can be summarized as follows.

Let $k$, $k \geq 2$, optimization iterations be performed, and the search information $A_k$ from (\ref{eq:13}) is obtained. To adaptively select the points of the next iterations, the optimization algorithm has to evaluate the possibility of placing the global minimum in the subintervals, into which the interval $[0,1]$ is divided by the points $x_i$, $1 \leq i \leq k$, previously performed iterations of the global search. This assessment can be performed by introducing characteristics of the subintervals, the value of which would be proportional to the degree to which the global minimum can be located in these subintervals

\begin{equation}\label{eq:15}
R(i)=\mathcal{R} (i,A_k ), 1 < i \leq k.
\end{equation}

The type of characteristics $R(i)$, $1 < i \leq k$, is a key design element of global optimization algorithms. For example, for an algorithm that builds a uniform dense grid in the search domain, the characteristics may simply be the length of the interval

\begin{equation}\label{eq:16}
R(i)=(x_i-x_{i-1}), 1 < i \leq k, \; x_i, x_{i-1} \; \mathrm{from} \; A_k.
\end{equation}

For the multidimensional generalizations of the algorithms proposed in \cite{c22,c23}, the characteristic is an estimate of the minimum possible value of the minimized function on the interval, i.e.

\begin{equation}\label{eq:17}
R(i)=0.5 H \rho_i - 0.5 (z_{i-1} + z_i), \rho_i=(x_i - x_{i-1})^{1/N}, 1 < i \leq k,
\end{equation}
where $H$ is the H\"older constant from (\ref{eq:12}) for the reduced global optimization problem to be solved; $N$ is the dimensionality of the problem from (\ref{eq:7}). For the multidimensional generalized global search algorithm (GSA) \cite{c5,c24}, developed in the framework of the information-statistical approach, the characteristic has the form

\begin{equation}\label{eq:18}
R(i)=m \rho_i+\frac{(z_i-z_{i-1})^2}{m \rho_i }-2(z_{i-1}+z_i ), \rho_i=(x_i-x_{i-1})^{1/N}  ,1 < i \leq k,
\end{equation}
where $m$ is the numerical estimate of the H\"older constant obtained from the available search information $A_k$ from (\ref{eq:13})

\begin{equation}\label{eq:19}
m=r M, M=\frac{\max|z_i-z_{i-1}|}{\rho_i}, \rho_i=(x_i-x_{i-1})^{1/N}  ,1 < i \leq k
\end{equation}
($r>1$ is the parameter of the GSA algorithm).

The presence of interval characteristics makes it possible to represent the procedure for iterating a global search as the following sequence of steps \cite{c5}.

\textit{Step 1.} Calculate the characteristics of the intervals $R(i)$, $1 < i \leq k$, and determine the interval with the maximum characteristic
\begin{equation}\label{eq:20}
R(t)=\max{R(i)}, 1 < i \leq k.
\end{equation}

\textit{Step 2.} Select the next iteration point in the interval with the maximum characteristic \footnote{The rule $X$ for choosing the point of the next iteration in the interval $(x_{t-1},x_t)$ is another design element of characteristically represented algorithms.}
\begin{equation}\label{eq:21}
x^{k+1}=X(x_{t-1},x_t ).
\end{equation}

Calculate the value $z^{k+1}$ of the minimized function at this point (the procedure for calculating the value of the function will be called a trial) and add the obtained data $(x^{k+1},z^{k+1})$ to the search information $A_k$ from (\ref{eq:13}).

\textit{Step 3.} Check the stopping condition 
\begin{equation}\label{eq:22}
\rho_t \leq \varepsilon,	
\end{equation}
where $\rho_t=(x_t-x_{t-1})^{1/N}$, $t$ from (\ref{eq:20}) and $\varepsilon > 0$ is the specified accuracy of the solution to the optimization problem. If the stopping condition (\ref{eq:22}) is reached, then the solution to the optimization problem is stopped, otherwise $k=k+1$ is assumed and the next iteration of the global search is performed.

After the calculations are completed, the lowest calculated value of the minimized function can be taken as the global minimum estimate
\begin{equation}\label{eq:23}
z_k^*=\min{z_i}, 1 \leq i \leq k.
\end{equation}

Again, we note that the above computational scheme is quite general. Within the framework of this scheme, many global search algorithms can be presented, as evidenced, in particular, by the examples given in (\ref{eq:16})--(\ref{eq:18}), as well as other characteristically represented algorithms (see, for example, \cite{c22,c23,c24,c25,c26,c27,c28,c29,c30,c31}). Thus, the parallelization methods discussed in Section \ref{sec:4} can be used for a sufficiently wide variety of global search algorithms.

The convergence conditions for the characteristically represented algorithms depend on the properties of the interval characteristics used. For example, one of the sufficient conditions for convergence of the algorithms is the requirement that, when iterating the global search, the characteristic of the interval containing the global minimum point repeatedly takes the maximum value in Step 1 of the general characteristic scheme. This condition is fulfilled, for example, for the multidimensional generalizations of the algorithms proposed in \cite{c22,c23} for the exact setting of the H\"older constant in (\ref{eq:17}). For GSA, the sufficient condition for convergence is the relation \cite{c5}
\begin{equation}\label{eq:24}
m \geq 2H,
\end{equation}
starting with some iteration of the global search. Additional information on the convergence of characteristically represented global search algorithms can be obtained, for example, in \cite{c5}.

In the framework of the proposed approach, the GSA algorithm described by rules (\ref{eq:18})-(\ref{eq:23}) will be used to solve global optimization problems of the set $F$. Moreover, due to the informational connectivity (\ref{eq:5}) of the problems of the set $F$, the GSA algorithm can use all the search information $A_k$ from (\ref{eq:13}) when solving the next global optimization problem. The GSA algorithm for solving problems of the set $F$ and using search information $A_k$, will be referred to as the multiple global search algorithm (MGSA).

@@

In keeping with the initial assumption that it is computationally expensive to calculate the values of minimized functions of the set $F$ from (\ref{eq:3}), solving all global optimization problems (\ref{eq:1}) may require a huge amount of calculations, which increases with the number of problems to be solved. Thus, obtaining estimates of globally optimal decisions with reasonable time delays can only be achieved using parallel high-performance computational systems.

>>>>

\usepackage{adjustbox}
\usepackage{amsmath}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
%\usepackage{algorithmic}
%\usepackage{algorithm}
\usepackage{algorithm,algpseudocode,float}

\algnewcommand\algorithmicforeach{\textbf{for each}}
\algdef{S}[FOR]{ForEach}[1]{\algorithmicforeach\ #1\ \algorithmicdo}

@@

@@

The first and obvious approach is to use local methods that allow one to efficiently solve optimization problems with unimodal objective functions \cite{c1}. In the case of multiextremal problems, local methods should be applied repeatedly to find global minima using different initial points. Research in this direction began as early as the 1960s and is still relevant today \cite{c10,c11,c12}. Unfortunately, the problem of choosing initial points still remains open -- even using a significant number of initial points does not guarantee a reliable solution of global optimization problems. Another obvious approach is to construct in the search domain  a number of grids in whose nodes values of functions to be minimized are computed \cite{c2,c3,c4,c5}. To reduce computational complexity of uniform grids, this approach envisages different methods for construction of non-uniform grids, which would be denser only in the vicinity of the global minimum. This line of research is developing in the field of so-called nature-inspired metaheuristic methods, including, for example, genetic algorithms, bee algorithms, particle swarm optimization, simulated annealing, ant colony optimization, and firefly algorithms, etc. \cite{c4,c13,c14,c15}. A large number of applied problems have been solved using these methods, however, as before, it is also difficult to obtain guaranteed estimates of the accuracy when finding globally-optimal solutions for these algorithms.

Yet another line of research in the field of global optimization is based on the assumption that the analytical form of the functions to be optimized is known. In this case, we can speak of d.c.(difference of convex) optimization problems in which nonconvex multiextremal functions are represented as a difference of two convex functions \cite{c16,c17}. Another approach to the development of global optimization methods in this area is based on the use of interval analysis methods \cite{c18}.

In case analytical expressions for the functions to be optimized are not known (and this is a typical situation for choosing the best solutions in applied optimization problems), it should be understood that numerical estimates of global optimal solutions cannot in principle be obtained if there are no a priori assumptions about the behaviour of the functions to be minimized. One if the most widely used ways for specifying such information is to assume the validity of the Lipschitz condition according to which possible changes of the function to be minimized are restricted by the variation values of the parameters being varied. The use of this assumption leads to a class of Lipschitzian global optimization problems that has been extensively explored \cite{c3,c6,c7,c19,c19,c20,c21,c22,c23}.  As a rule, algorithms for solving this class of problems are adaptive, i.e. when determining the points for global search iterations, algorithms tend to use all the search information obtained in the process of computation. Such adaptivity leads to non-uniform grids and makes it possible to construct estimates of the global minimum based on a finite number of global search iterations. A large number of results in this direction have been obtained within the framework of the rapidly developing information-statistical theory of global optimization \cite{c6}. It should also be noted that multivariate algorithms of this class quite often use dimensionality reduction using Peano curves or nested optimization scheme \cite{c23}. One can also mention the class of DIRECT-type methods using several estimates of the Lipschitz constant chosen from some set of possible values; a detailed review of this line of research is given in \cite{c22}.

The paper proposes a further complication in formulating the decision making problems, when not only one, but several global optimization problems need to be solved simultaneously. Such problems are encountered, for example, when one needs to find several efficient solutions of multicriteria optimization problems or when part of the variable parameters in the optimization problem can take only a certain number of discrete values. For the first time, the formulation of such multiple global optimization (MGO was considered in \cite{c25,c26}, and then was investigated in \cite{c27,c28}. The relatively limited attention given to MGO problems is most likely due to the high computational complexity involved for solving even one GO problem.

In general, MGO problems can be solved by a simple sequential (or parallel) solution of individual problems from a set. However, this approach results in a many-fold increase in the computational complexity of solving such problems. Meanwhile, MGO problems may have a certain informational interrelation -- a similar connectivity is observed, for example, when methods based on scalarization of the vector efficiency criterion are applied to solve multicriteria optimization (MCO) problems. In this case, it becomes necessary to solve multiple scalar (in the general case, global) optimization problems in order to find several efficient decisions  for MCO problems -- see, for example, \cite{c29,c30,c31,c32,c33}. Such interconnectedness of MGO problems can be used to reduce significantly the amount of computation performed when solving the entire set of problems.

The approach proposed in this paper assumes that the computational values of any function of the set of MGO problems can be recalculated to the values of any other functions of the set without any time consuming calculations. In such cases, all the search information obtained in solving a particular problem can be used in solving all other problems in the set. Such repeated reuse of search information can significantly reduce the amount of computation performed, up to just a few iterations when solving the next MGO problems. In addition, the presence of informational connectivity significantly expands the possibilities of a parallel solution for MGO problems. The paper proposes two new parallel computational schemes: a cooperative scheme, which means the search information between the problems is exchanged when solving several MGO problems in parallel, and a competitive scheme, which occurs when MGO problems solved in parallel compete with each other for the use of available computational resources.

The structure of the paper is as follows. Section \ref{sec:2} sets out the problem of multiple global optimization. Section \ref{sec:3} presents a general computational scheme for solving global search problems. Section \ref{sec:4} proposes parallel methods for solving MGO problems by using computational systems with shared and distributed memory. Section \ref{sec:5} contains the results of numerical experiments confirming the effectiveness of the proposed parallel methods. In the conclusion, results are discussed and the main directions for further research are presented.

@@

\begin{equation}\label{eq:2}
D  = \{ y\in R^N: a_i \leq y_i \leq b_i, 1 \leq i \leq N \}
\end{equation}
defined by a set of optimized functions
\begin{equation}\label{eq:3}
F(y) = \{ f_1(y),  f_2(y),\dots, f_s(y) \},
\end{equation}
where $y = (y_1,y_2,\dots,y_N)$ is the vector of varied parameters, $N$ is the dimension of the global optimization problems to be solved, and $D$ is the search domain for the given vectors $a$ and $b$.

In this paper, it is assumed that the functions $f_i(y)$, $1 \leq i \leq s$, can be multiextremal, and the procedures for calculating the values of these functions at points in the search domain $y \in D$ can be computationally expensive. It is also assumed that the functions $f_i(y)$, $1 \leq i \leq s$, satisfy the Lipschitz condition
\begin{equation}\label{eq:4}
|f_i (y')-f_i (y'')| \leq L_i \|y'-y''\|, y',y''\in D, 1 \leq i \leq s,
\end{equation}
where $L_i$ is the Lipschitz constant for the function $f_i(y)$, $1 \leq i \leq s$, and ${\|*\|}$ denotes the Euclidean norm in $R^N$. Fulfillment of the Lipschitz condition allows us to construct numerical estimates of the possible behavior of minimized functions based on a finite set of calculated function values at points in the search domain.

A key condition of this approach is the assumption that any calculated value of any function $f_i(y)$, $1 \leq i \leq s$, of the set F at an arbitrary point $y \in D$ of the search domain $D$ can be transformed to the value of any other function of the set $F$ at the same point $y \in D$ without any time consuming computations, in a time significantly shorter compared to the computation time values of the functions $f_i(y)$, $1 \leq i \leq s$. In other words, there must be a transformation $Z$ such that
\begin{equation}\label{eq:5}
\begin{matrix}
\forall 1 \leq i, j \leq s, \bar{y} \in D \Rightarrow z_i=Z(z_j ), z_i = f_i (\bar{y}),z_j=f_j (\bar{y}), \\
T(Z) \ll T(f_l (y) ),1 \leq l \leq s,
\end{matrix}
\end{equation}
where $T(*)$  is the computation time of the expression given in parenthesis.

This condition is fulfilled, for example, when solving multicriteria optimization (MCO) problems 
\begin{equation}\label{eq:6}
g(y) = (g_1(y), g_2(y), \dots , g_s(y)) \to min,  y\in D.
\end{equation}

To find efficient (unimprovable) solutions, a widely used approach is based on the reduction (scalarization) of the vector criterion $g(y)$ to the general scalar criterion, the minimization of which will lead to the Pareto optimal solution of the MCO problem \cite{c29,c30,c31,c32,c33}. So, when using mini-max convolution, the vector criterion $g(y)$ from (\ref{eq:6}) reduces to the scalar criterion
\begin{equation}\label{eq:7}
\begin{matrix}
G(\lambda,y)=\max{(\lambda_i g_i(y),1 \leq i \leq s)},	\\
\lambda=(\lambda_1,\lambda_2,\dots,\lambda_s)\in \Lambda \subset R^s:\sum_{i=1}^s{\lambda_i =1},\lambda_i\geq 0,1 \leq i \leq s.
\end{matrix}
\end{equation}

If it is necessary to find several efficient decisions to the MCO problem (\ref{eq:6}), the optimization of the scalar criterion $G(\lambda,y)$ from (\ref{eq:7}) is performed for different values of the coefficients $\lambda \in \Lambda$. Thus, the problem of finding a set of efficient decisions for the MCO problem can be represented as a statement of the MGO problem (\ref{eq:1})-(\ref{eq:3}), where the optimized functions $f_i(y)$, $1 \leq i \leq s$, and the set $F$ are defined as
\begin{equation}\label{eq:8}
f_i (y)= G(\lambda_i,y),1 \leq i \leq s.
\end{equation}

As a result, if the values of efficiency criteria $g_i(y)$, $1 \leq i \leq s$, are computed at some point $y=\bar{y}$ of the search domain $D$, then the value of the scalar criterion $G(\lambda,y)$ can be computed at the point $\bar{y} \in D$ for any values of the coefficients $\lambda \in \Lambda$. In this case the transformation of $Z$ from (\ref{eq:5}) is determined by the expression (\ref{eq:7}) at fixed values of the criteria $g_l (y)$,$1 \leq l \leq m$) \cite{c27,c28}.

The ability to quickly convert the calculated values of optimized functions determines the need to accumulate and use search information obtained in the calculation process
\begin{equation}\label{eq:9}
\omega_k (j)=\{(y^i,f^i=f_j(y^i)):1 \leq i \leq k \},1 \leq j \leq s,
\end{equation}
where $j$, $1 \leq j \leq s$, is the number of the current function being minimized $f_j (y)$ of the set $F$ from (\ref{eq:3}), k is the number of iterations of the global search, $y^i \in D$, $1 \leq i \leq k$,  are points in which the values of the optimized problems of the set $F$ were computed, and $f^i$, $1 \leq i \leq k$, are the values of the function $f_j (y)$ at these points (the procedure for calculating the value of the function will be called a trial). Global optimization methods can use this search information, thereby ensuring the ability to solve each successive problem of the $F$ set, taking into account the results of all previous calculations. Numerical experiments that were performed (see Section \ref{sec:5}), confirm that such repeated use of search information reduces the number of iterations of the global search required to solve the next problem of the set $F$ (even to the point of performing just a few search iterations).

In conclusion, it should be noted that, in general, the set of functions being minimized $F$ from (\ref{eq:3}) can be formed dynamically in the course of calculations by adding new or removing existing global optimization functions
\begin{equation}\label{eq:10}
F'=F+(-) f(y).
\end{equation}

@@

The approach proposed in this paper is based on two key ideas used in the framework of the information-statistical theory of global optimization \cite{c6}.

In accordance with this theory, first of all, dimensionality reduction is achieved using the Peano space-filling curves (evolvents) $y(x)$, uniquely and continuously mapping the interval [0,1] to the $N$-dimensional domain $D$ -- see, for example, \cite{c6,c23}. As a result of this reduction, multidimensional global optimization problems (\ref{eq:1})-(\ref{eq:2}) can be reduced to one-dimensional ones:

@@

Dimensionality reduction significantly decreases the computational complexity of analyzing multidimensional data to highlight the subdomains of the search domain $D$ for the efficient continuation of the global search, and reduces the severity of the ``curse of dimensionality'' problem in solving multidimensional optimization problems. After dimensionality reduction, the search information $\omega_k$ from (\ref{eq:9}) takes the form
\begin{equation}\label{eq:13}
A_k (j)=\{ (x_i, y(x_i), z_i=f_j (y(x_i) ): 1 \leq i \leq k,1 \leq j \leq s \},
\end{equation}
in which, in addition to the reduced representation, the data is arranged in the ascending order\footnote{Data ordering is reflected by using a subscript.} of points $x_i$,  $1 \leq i \leq k$, i.e.

\begin{equation}\label{eq:14}
x_1< x_2< \cdots < x_k
\end{equation}
or a more efficient execution of global search algorithms. It should be noted that in optimizing the reduced one-dimensional functions $f_i (y(x))$, $1 \leq i \leq s$, many one-dimensional global search algorithms can be used (possibly after some generalization).

Numerical methods for constructing approximations of Peano space-filling curves (evolvents) with a given accuracy for a given dimensionality are considered in \cite{c6}; examples of such approximations are shown in Fig.\ref{fig:1}.

Further, in the framework of information-statistical theory, a general computational scheme of global optimization algorithms is proposed \cite{c6}. This scheme can be summarized as follows.

@@

Let $f(y)$ be the function to be minimized for which $k$, $k \geq 2$, optimization iterations are performed, and the search information $A_k$ from (\ref{eq:13}) is obtained. To adaptively select the points of the next iterations, the optimization algorithm has to evaluate the possibility of placing the global minimum in the subintervals, into which the interval [0,1] is divided by the points $x_i$, $1 \leq i \leq k$, of previously performed iterations of the global search. This assessment can be performed by introducing \textit{characteristics} of the subintervals, the value of which would be proportional to the degree to which the global minimum can be located in these subintervals
\begin{equation}\label{eq:15}
R(i)=\mathcal{R} (i,A_k ), 1 < i \leq k.
\end{equation}

The type of characteristics $R(i)$, $1 < i \leq k$, is a key design element of global optimization algorithms. For example, for an algorithm that builds a uniform dense grid in the search domain, the characteristics may simply be the length of the interval
\begin{equation}\label{eq:16}
R(i)=(x_i-x_{i-1}), 1 < i \leq k,
\end{equation}
where the points $x_i$, $1 < i \leq k$, are from $A_k$.

For the multidimensional generalizations of the algorithms proposed in \cite{c34,c35}, the characteristic is an estimate of the minimum possible value of the function being minimized on the interval, i.e.
\begin{equation}\label{eq:17}
R(i)=0.5 H \rho_i - 0.5 (z_{i-1} + z_i), \rho_i=(x_i - x_{i-1})^{1/N}, 1 < i \leq k,
\end{equation}
where $H$ the values $z_i$, $1<i \leq k$, are from $A_k$, H is the H\"older constant from (\ref{eq:12}) for the reduced global optimization problem to be solved and $N$ is the dimensionality of the problem from (\ref{eq:1}). For the multidimensional generalized global search algorithm (GSA) \cite{c6,c23}, developed in the framework of the information-statistical approach, the characteristic has the form
\begin{equation}\label{eq:18}
R(i)=m \rho_i+\frac{(z_i-z_{i-1})^2}{m \rho_i }-2(z_{i-1}+z_i ), \rho_i=(x_i-x_{i-1})^{1/N}  ,1 < i \leq k,
\end{equation}
where $m$ is the numerical estimate of the H\"older constant obtained from the available search information $A_k$ from (\ref{eq:13})

\begin{equation}\label{eq:19}
m=r M, M=\frac{\max|z_i-z_{i-1}|}{\rho_i}, \rho_i=(x_i-x_{i-1})^{1/N}  ,1 < i \leq k
\end{equation}
($r>1$ is the reliability parameter of the GSA algorithm).

A detailed description of the GSA algorithm is given in Algorithm \ref{alg:1}(see also \cite{c6}). The inputs of the algorithm are the function to be minimized $f(y)$\footnote{ }, the search domain $D$ from (\ref{eq:2}), the reliability parameter $r > 1$, the required accuracy $\varepsilon > 0$ and the maximum permissible number of global search iterations $kmax$. The output of the algorithm is an estimate of the global minimum point $y_k^* =y(x_k^*)$, the value of the function to be minimized at this point $z_k^* =f(y_k^*)$  and the number of performed global search iterations $k$. In the course of calculations, the search information $A_k$\footnote{ } from (\ref{eq:13}) is used, which contains the points of previously performed global search iterations $x_i$, $1 \leq i \leq k$, ordered in ascending order, and the values of the function to be minimized $z_i=f(y(x_i))$, $1\leq i \leq k$.

\begin{algorithm}[]
\caption{PSEUDO CODE OF THE GSA ALGORITHM} \label{alg:1}
\scriptsize
\hspace*{\algorithmicindent} \textbf{Input}: $f(y)$, $D$, $r$, $\varepsilon$, $K$;\\
\hspace*{\algorithmicindent} \textbf{Output}: $y_k^*$, $z_k^*$, $k$;
\begin{algorithmic}[1]
\State Initialize $A_2=\{(x_1=0,y_1=y(x_1 ),z_1=f(y(x_1 )),(x_2=1,y_2=y(x_2 ),z_2=f(y(x_2 ))\}$, $k = 2$, $stop = false$;
\While{$k < kmax$ \textbf{and not} $stop$} 
  \State Calculate the estimate of the H\"older constant $m$ from (\ref{eq:19}) using $A_k$;
  \State Calculate the interval characteristics $R(i)$ from (\ref{eq:18}) \textbf{for all} $i = 2, \dots, k$;
	\State Find the maximum characteristic 
	\begin{equation}\label{eq:20}
		R(t)=\max{R(i)}, 1 < i \leq k.
  \end{equation}
	\State Calculate the next iteration point
	\begin{equation}\label{eq:21}
	\begin{matrix}
	   x^{k+1}=\frac{x_t+x_{t-1}}{2}-sign(z_t-z_{t-1})(\frac{|z_t-z_{t-1}|}{m})^N/2r \\
		 sign(v)=\begin{cases}
		            1, &v>0,\\
		            0, &v=0,\\
								-1,&v<0
							\end{cases}
	\end{matrix}
	\end{equation}

	\State
	\begin{equation}\label{eq:22}
     stop = (x_t - x_{t-1})^{1/N} \leq \varepsilon.
  \end{equation}
  \If{ \textbf{not} $stop$}
	  \State Calculate $z^{k+1} = f( y(x^{k+1}))$;
    \State Insert $(x^{k+1}, y(x^{k+1}), z^{k+1})$ into the search information $A_k$;
    \State Increase $k = k + 1$;
  \EndIf
\EndWhile
\State Find the estimate of the global minimum point
\begin{equation}\label{eq:23}
y_k^*= y(x_l),  z_k^*=z_l,
\end{equation}
where $l= \argmin_{1 \leq i \leq k}z_i$.
\State \textbf{return} $y_k^*$, $z_k^*$, $k$;
\end{algorithmic}
\end{algorithm}

The sufficient condition for convergence of the GSA algorithm is the relation
\begin{equation}\label{eq:24}
m \geq 2H,
\end{equation}
that should be satisfied starting with some iteration of the global search.

In the framework of the proposed approach, the GSA algorithm described above will be used to solve global optimization problems of the set $F$. Moreover, due to the informational connectivity (\ref{eq:5}) of the problems of the set $F$, the search information $A_k$ can be quickly transformed to the values of the next function to be minimized $f(y)$ of the set $F$ and GSA can minimize each successive function $f(y)$, taking into account the results of all previous calculations. The GSA algorithm for solving problems of the set $F$ and using search information $A_k$, will be referred to as the multiple global search algorithm (MGSA).

@@

In keeping with the initial assumption that it is computationally expensive to calculate the values of the functions to be minimized of the set $F$ from (\ref{eq:3}), solving all global optimization problems (\ref{eq:1}) may require a huge amount of calculations, which increases with the number of problems to be solved. Thus, obtaining estimates of globally optimal decisions with reasonable time delays can only be achieved using parallel high-performance computational systems.
